---
title: "Chapter 11: Difference-in-Differences and Event Studies"
author: "Junlong Feng"
format:
  live-html:
    toc: true
    toc-depth: 2
execute:
  enabled: false
engine: markdown
---

## Simulated Staggered-Adoption Panel

Generates treatment cohorts with effects that vary by cohort and exposure length.

```{r}
set.seed(5280)
N <- 600
TT <- 10

units <- data.frame(
  id = 1:N,
  G = sample(c(0, 4, 6, 8), N, replace = TRUE,
             prob = c(0.30, 0.25, 0.25, 0.20)),
  alpha = rnorm(N),
  W = rnorm(N)
)

data <- merge(expand.grid(id = 1:N, t = 1:TT), units,
              by = "id", sort = TRUE)
data$D <- as.integer(data$G > 0 & data$t >= data$G)
data$event_time <- ifelse(data$G > 0, data$t - data$G, NA)

data$cohort_scale <- ifelse(data$G == 4, 1.50,
                     ifelse(data$G == 6, 1.00,
                     ifelse(data$G == 8, 0.60, 0)))
data$tau <- ifelse(data$D == 1,
                   data$cohort_scale * (1 + 0.5 * data$event_time), 0)

data$lambda <- 0.25 * data$t + 0.03 * data$t^2
data$Y0 <- data$alpha + data$lambda + 0.5 * data$W +
           rnorm(nrow(data))
data$Y <- data$Y0 + data$tau
```

## One Group-Time DID by Hand

Calculates the group-4, period-6 DID and compares it with the known simulated effect.

```{r}
simple_att_gt <- function(data, g, t) {
  stopifnot(t >= g)
  base <- g - 1

  now <- data[data$t == t, c("id", "G", "Y")]
  names(now)[names(now) == "Y"] <- "Y_now"
  before <- data[data$t == base, c("id", "Y")]
  names(before)[names(before) == "Y"] <- "Y_before"

  change <- merge(now, before, by = "id")
  change$dY <- change$Y_now - change$Y_before
  treated <- change$G == g
  controls <- change$G == 0 | change$G > t

  mean(change$dY[treated]) - mean(change$dY[controls])
}

simple_att_gt(data, g = 4, t = 6)
mean(data$tau[data$G == 4 & data$t == 6])
```

## Callaway-Sant'Anna Group-Time Effects

Estimates and plots group-time effects with bootstrap confidence bands.

```{r}
library(did)

att <- att_gt(
  yname = "Y", tname = "t", idname = "id", gname = "G",
  xformla = ~ 1, data = data, panel = TRUE,
  control_group = "notyettreated", est_method = "dr",
  base_period = "universal", clustervars = "id",
  bstrap = TRUE, cband = TRUE,
  biters = 199                 # short classroom demonstration only
)

summary(att)
ggdid(att)
```

## Aggregate DID Effects

Reports overall, cohort, dynamic, and balanced dynamic aggregations.

```{r}
overall <- aggte(att, type = "simple")
by_group <- aggte(att, type = "group")
dynamic <- aggte(att, type = "dynamic", min_e = -3, max_e = 4)

summary(overall)
summary(by_group)
summary(dynamic)
ggdid(dynamic)

dynamic_balanced <- aggte(
  att, type = "dynamic",
  min_e = -3, max_e = 3,
  balance_e = 3
)
summary(dynamic_balanced)
ggdid(dynamic_balanced)
```

## TWFE versus Sun-Abraham Event Studies

Fits and plots conventional TWFE and interaction-weighted event-study estimates.

```{r}
library(fixest)

data$ever_treated <- as.integer(data$G > 0)
data$event_twfe <- ifelse(data$G > 0, data$t - data$G, 0)

twfe_event <- feols(
  Y ~ i(event_twfe, ever_treated, ref = -1) | id + t,
  data = data, cluster = ~ id
)

# A cohort outside the observed period range is treated as never treated.
data$G_sunab <- ifelse(data$G == 0, 1000, data$G)
sa_event <- feols(
  Y ~ sunab(G_sunab, t, ref.p = -1) | id + t,
  data = data, cluster = ~ id
)

iplot(list("Conventional TWFE" = twfe_event,
           "Sun--Abraham" = sa_event),
      ref.line = -1, main = "Event-study estimates")
```
