---
title: "Chapter 11: Difference-in-Differences and Event Studies"
author: "Junlong Feng"
format:
  live-html:
    toc: true
    toc-depth: 2
execute:
  enabled: false
engine: markdown
---

## Simulated Staggered-Adoption Panel

Generates persistent, temporary, and delayed treatment effects across cohorts.

```{r}
set.seed(5280)
N <- 600
TT <- 10

units <- data.frame(
  id = 1:N,
  G = sample(c(0, 4, 6, 8), N, replace = TRUE,
             prob = c(0.30, 0.25, 0.25, 0.20)),
  alpha = rnorm(N),
  W = rnorm(N)
)

data <- merge(expand.grid(id = 1:N, t = 1:TT), units,
              by = "id", sort = TRUE)
data$D <- as.integer(data$G > 0 & data$t >= data$G)
data$event_time <- ifelse(data$G > 0, data$t - data$G, NA)

# Strongly heterogeneous treatment-effect paths:
# G = 4: persistent; G = 6: temporary; G = 8: delayed.
data$tau <- 0
data$tau[data$G == 4 & data$D == 1] <- 10
data$tau[data$G == 6 & data$D == 1 & data$event_time <= 1] <- 10
data$tau[data$G == 8 & data$D == 1 & data$event_time >= 2] <- 10

data$lambda <- 0.25 * data$t + 0.03 * data$t^2
data$Y0 <- data$alpha + data$lambda + 0.5 * data$W +
           rnorm(nrow(data))
data$Y <- data$Y0 + data$tau

# Known effects from the data-generating process.
true_gt <- aggregate(
  tau ~ G + t + event_time,
  data = data[data$G > 0, ],
  FUN = mean
)
names(true_gt)[names(true_gt) == "tau"] <- "true_effect"

true_event <- aggregate(
  tau ~ event_time,
  data = data[data$G > 0, ],
  FUN = mean
)
names(true_event)[names(true_event) == "tau"] <- "true_effect"

cohort_n <- table(units$G)
true_effects_by_cohort <- true_gt[
  true_gt$t >= true_gt$G,
  c("G", "event_time", "true_effect")
]
print(true_effects_by_cohort, digits = 4, row.names = FALSE)
```

## One Group-Time DID by Hand

Calculates the group-4, period-6 DID and compares it with the known simulated effect.

```{r}
simple_att_gt <- function(data, g, t) {
  stopifnot(t >= g)
  base <- g - 1

  now <- data[data$t == t, c("id", "G", "Y")]
  names(now)[names(now) == "Y"] <- "Y_now"
  before <- data[data$t == base, c("id", "Y")]
  names(before)[names(before) == "Y"] <- "Y_before"

  change <- merge(now, before, by = "id")
  change$dY <- change$Y_now - change$Y_before
  treated <- change$G == g
  controls <- change$G == 0 | change$G > t

  mean(change$dY[treated]) - mean(change$dY[controls])
}

hand_comparison <- data.frame(
  estimand = "ATT(4,6)",
  estimate = simple_att_gt(data, g = 4, t = 6),
  true_effect = mean(data$tau[data$G == 4 & data$t == 6])
)
hand_comparison$estimation_error <-
  hand_comparison$estimate - hand_comparison$true_effect
print(hand_comparison, digits = 4, row.names = FALSE)
```

## Callaway-Sant'Anna Group-Time Effects

Estimates and plots group-time effects with bootstrap confidence bands.

```{r}
library(did)

att <- att_gt(
  yname = "Y", tname = "t", idname = "id", gname = "G",
  xformla = ~ 1, data = data, panel = TRUE,
  control_group = "notyettreated", est_method = "dr",
  base_period = "universal", clustervars = "id",
  bstrap = TRUE, cband = TRUE,
  biters = 199                 # short classroom demonstration only
)

summary(att)

att_truth <- data.frame(
  group = att$group,
  period = att$t,
  event_time = att$t - att$group,
  estimate = att$att
)
att_truth$true_effect <- true_gt$true_effect[
  match(
    paste(att_truth$group, att_truth$period),
    paste(true_gt$G, true_gt$t)
  )
]
att_truth$estimation_error <- att_truth$estimate - att_truth$true_effect
print(att_truth, digits = 4, row.names = FALSE)

ggdid(att)
```

## Aggregate DID Effects

Reports overall, cohort, dynamic, and balanced dynamic aggregations.

```{r}
overall <- aggte(att, type = "simple")
by_group <- aggte(att, type = "group")
dynamic <- aggte(att, type = "dynamic", min_e = -3, max_e = 4)

true_group <- aggregate(
  tau ~ G,
  data = data[data$D == 1, ],
  FUN = mean
)
names(true_group)[names(true_group) == "tau"] <- "true_effect"
true_group$weight <- as.numeric(
  cohort_n[as.character(true_group$G)]
)

simple_comparison <- data.frame(
  estimand = "Simple overall ATT",
  estimate = overall$overall.att,
  true_effect = mean(data$tau[data$D == 1])
)
simple_comparison$estimation_error <-
  simple_comparison$estimate - simple_comparison$true_effect

group_comparison <- data.frame(
  group = by_group$egt,
  estimate = by_group$att.egt
)
group_comparison$true_effect <- true_group$true_effect[
  match(group_comparison$group, true_group$G)
]
group_comparison$estimation_error <-
  group_comparison$estimate - group_comparison$true_effect

group_overall_comparison <- data.frame(
  estimand = "Group-weighted overall ATT",
  estimate = by_group$overall.att,
  true_effect = weighted.mean(true_group$true_effect, true_group$weight)
)
group_overall_comparison$estimation_error <-
  group_overall_comparison$estimate - group_overall_comparison$true_effect

dynamic_comparison <- data.frame(
  event_time = dynamic$egt,
  estimate = dynamic$att.egt
)
dynamic_comparison$true_effect <- true_event$true_effect[
  match(dynamic_comparison$event_time, true_event$event_time)
]
dynamic_comparison$estimation_error <-
  dynamic_comparison$estimate - dynamic_comparison$true_effect

dynamic_overall_comparison <- data.frame(
  estimand = "Dynamic overall ATT",
  estimate = dynamic$overall.att,
  true_effect = mean(dynamic_comparison$true_effect[
    dynamic_comparison$event_time >= 0
  ])
)
dynamic_overall_comparison$estimation_error <-
  dynamic_overall_comparison$estimate -
  dynamic_overall_comparison$true_effect

summary(overall)
print(simple_comparison, digits = 4, row.names = FALSE)

summary(by_group)
print(group_comparison, digits = 4, row.names = FALSE)
print(group_overall_comparison, digits = 4, row.names = FALSE)

summary(dynamic)
print(dynamic_comparison, digits = 4, row.names = FALSE)
print(dynamic_overall_comparison, digits = 4, row.names = FALSE)
ggdid(dynamic)

dynamic_balanced <- aggte(
  att, type = "dynamic",
  min_e = -3, max_e = 3,
  balance_e = 3
)

balanced_groups <- sort(unique(
  data$G[data$G > 0 & TT - data$G >= 3]
))
true_event_balanced <- aggregate(
  tau ~ event_time,
  data = data[data$G %in% balanced_groups, ],
  FUN = mean
)
names(true_event_balanced)[names(true_event_balanced) == "tau"] <-
  "true_effect"

dynamic_balanced_comparison <- data.frame(
  event_time = dynamic_balanced$egt,
  estimate = dynamic_balanced$att.egt
)
dynamic_balanced_comparison$true_effect <-
  true_event_balanced$true_effect[
    match(
      dynamic_balanced_comparison$event_time,
      true_event_balanced$event_time
    )
  ]
dynamic_balanced_comparison$estimation_error <-
  dynamic_balanced_comparison$estimate -
  dynamic_balanced_comparison$true_effect

dynamic_balanced_overall_comparison <- data.frame(
  estimand = "Balanced dynamic overall ATT",
  estimate = dynamic_balanced$overall.att,
  true_effect = mean(dynamic_balanced_comparison$true_effect[
    dynamic_balanced_comparison$event_time >= 0
  ])
)
dynamic_balanced_overall_comparison$estimation_error <-
  dynamic_balanced_overall_comparison$estimate -
  dynamic_balanced_overall_comparison$true_effect

summary(dynamic_balanced)
print(dynamic_balanced_comparison, digits = 4, row.names = FALSE)
print(dynamic_balanced_overall_comparison, digits = 4, row.names = FALSE)
ggdid(dynamic_balanced)
```

## TWFE versus Sun-Abraham Event Studies

Fits and plots conventional TWFE and interaction-weighted event-study estimates.

```{r}
library(fixest)

data$ever_treated <- as.integer(data$G > 0)
data$event_twfe <- ifelse(data$G > 0, data$t - data$G, 0)

twfe_event <- feols(
  Y ~ i(event_twfe, ever_treated, ref = -1) | id + t,
  data = data, cluster = ~ id
)

# A cohort outside the observed period range is treated as never treated.
data$G_sunab <- ifelse(data$G == 0, 1000, data$G)
sa_event <- feols(
  Y ~ sunab(G_sunab, t, ref.p = -1) | id + t,
  data = data, cluster = ~ id
)

event_study_truth <- true_event[order(true_event$event_time), ]
print(event_study_truth, digits = 4, row.names = FALSE)

event_models <- list(
  "Conventional TWFE" = twfe_event,
  "Sun-Abraham" = sa_event
)
plot_cols <- c("#0072B2", "#D55E00", "#009E73")
plot_pch <- c(16, 17, 15)

# Read the parameters chosen by iplot so its confidence limits include truth.
plot_info <- iplot(
  objects = event_models,
  only.params = TRUE,
  sep = 0.15
)
event_labels <- unique(suppressWarnings(
  as.numeric(as.character(plot_info$labels))
))
event_labels <- event_labels[is.finite(event_labels)]
truth_in_plot <- event_study_truth[
  event_study_truth$event_time %in% event_labels,
]
truth_in_plot <- truth_in_plot[order(truth_in_plot$event_time), ]

plot_ylim <- range(
  plot_info$prms$ci_low,
  plot_info$prms$ci_high,
  truth_in_plot$true_effect,
  na.rm = TRUE
)
plot_pad <- 0.05 * diff(plot_ylim)

iplot(
  objects = event_models,
  ref.line = -1,
  col = plot_cols[1:2],
  pch = plot_pch[1:2],
  pt.join = TRUE,
  sep = 0.15,
  ylim = plot_ylim + c(-plot_pad, plot_pad),
  xlab = "Event time",
  ylab = "Treatment effect",
  main = "TWFE and Sun-Abraham event-study estimates"
)

lines(
  truth_in_plot$event_time, truth_in_plot$true_effect,
  col = plot_cols[3], lty = 2, lwd = 2
)
points(
  truth_in_plot$event_time, truth_in_plot$true_effect,
  col = plot_cols[3], pch = plot_pch[3]
)
legend(
  "topleft",
  legend = c("Conventional TWFE", "Sun-Abraham", "True ATT(e)"),
  col = plot_cols,
  pch = plot_pch,
  lty = c(1, 1, 2),
  lwd = c(1, 1, 2),
  bty = "n"
)
```
